---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

```{r}
library(readr)
# Extract the numeric part from the column names for deaths
death_columns <- colnames(av)[grepl("^Death\\d+$", colnames(av))]
time_values <- parse_number(death_columns)

# Create the deaths dataset
deaths <- data.frame(Time = time_values, Death = character(length(time_values)), stringsAsFactors = FALSE)

# Populate the Death column based on the original dataset
for (i in 1:nrow(deaths)) {
  deaths$Death[i] <- ifelse(av[i, death_columns[i]] == "YES", "yes", "no")
}
```


Similarly, deal with the returns of characters.

```{r}
# Extract the numeric part from the column names for returns
return_columns <- colnames(av)[grepl("^Return\\d+$", colnames(av))]
time_values_return <- parse_number(return_columns)

# Create the returns dataset
returns <- data.frame(Time = time_values_return, Return = character(length(time_values_return)), stringsAsFactors = FALSE)

# Populate the Return column based on the original dataset
for (i in 1:nrow(returns)) {
  returns$Return[i] <- ifelse(av[i, return_columns[i]] == "YES", "yes", "no")
}
```


Based on these datasets calculate the average number of deaths an Avenger suffers. 

```{r}
# Calculate the average number of deaths
average_deaths <- mean(deaths$Time)
average_deaths
```

**On average, an Avenger suffers approximately 3 deaths**

## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Emily:\

Fact: The MVP of the Earth-616 Marvel Universe Avengers has to be Jocasta — an android based on Janet van Dyne and built by Ultron — who has been destroyed five times and then recovered five times.\

```{r}
library(dplyr)
# Filter the data for Jocasta
jocasta <- av %>%
  filter(Name.Alias == "Jocasta")

# Calculate the number of times Jocasta has been destroyed and recovered
num_deaths <- sum(jocasta$Death1 == "YES", jocasta$Death2 == "YES", jocasta$Death3 == "YES", jocasta$Death4 == "YES", jocasta$Death5 == "YES")
num_returns <- sum(jocasta$Return1 == "YES", jocasta$Return2 == "YES", jocasta$Return3 == "YES", jocasta$Return4 == "YES", jocasta$Return5 == "YES")

num_deaths
num_returns
```

The dataset showed that Jocasta indeed was destroyed 5 times and then recovered 5 times.\

Ryan:\

Fact: There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife, but only a 50 percent chance they recovered from a second or third death.\

```{r }
# Calculating ratio of those who return from their first death
firstdeath <- sum(av$Death1 == "YES")
firstreturn <- sum(av$Return1 == "YES")
firstreturn / firstdeath

# Calculating ratio of those who return from second/third death
stdeath <- sum(av$Death2 == "YES",
               av$Death3 == "YES")
streturn <- sum(av$Return2 == "YES",
                av$Return3 == "YES")
streturn / stdeath
```
    
## Individually - Mason
### FiveThirtyEight Statement

> Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team. That’s about 40 percent of all people who have ever signed on to the team.

### Include the code

```{r}
# Fact-checking the statement about the number of Avengers who died at least once
avengers_with_deaths <- av %>% 
  filter(Death1 == "YES")

number_of_avengers_with_deaths <- nrow(avengers_with_deaths)
number_of_avengers <- nrow(av)

number_of_avengers_with_deaths
number_of_avengers_with_deaths/number_of_avengers *100


```

### Include your answer

The number of avengers with at least 1 death was 69, which is accurate. Out of the total of 173 the percentage of deaths were 39.88%, but for this case you can't have a decimal percentage so it is rounded up to 40%, which makes this statement accurate.




